
## Required Packages
```{r}
require(prophet)
require(dplyr)
require(xts)
require(highcharter)
require(ggplot2)
require(tidyr)
require(viridis)
require(plotly)
require(lubridate)
require(data.table)
require(randomForest)
require(class)
```

## Preparing Dataset
```{r}
chic_crime_2008_2011 <- read.csv("D:/Study/Northeastern/Fall-2017/DM-ML/Homework/Project/Dataset/Chicago_Crimes_2008_to_2011.csv")
chic_crime_2012_2016 <- read.csv("D:/Study/Northeastern/Fall-2017/DM-ML/Homework/Project/Dataset/Chicago_Crimes_2012_to_2016.csv")
chic_crime_2012_2016<- chic_crime_2012_2016[!chic_crime_2012_2016$Year > 2016,]
chic_crime <- rbind(chic_crime_2008_2011, chic_crime_2012_2016)
```

## Creating data factors
```{r}
chic_crime$Date <- as.Date(chic_crime$Date, "%m/%d/%Y %I:%M:%S %p")
chic_crime$Day <- factor(day(as.POSIXlt(chic_crime$Date, format="%m/%d/%Y %I:%M:%S %p")))
chic_crime$Month <- factor(month(as.POSIXlt(chic_crime$Date, format="%m/%d/%Y %I:%M:%S %p")))
chic_crime$Year <- factor(year(as.POSIXlt(chic_crime$Date, format="%m/%d/%Y %I:%M:%S %p")))
chic_crime$Weekday <- factor(wday(as.POSIXlt(chic_crime$Date, format="%m/%d/%Y %I:%M:%S %p")))
```

## Grouping the dataset on various parameters
```{r}
by_Date <- (chic_crime) %>% group_by(Date) %>% summarise(Total = n())
by_Date<- na.omit(by_Date)
by_year <- chic_crime %>% group_by(Year) %>% summarise(Total = n()) %>% arrange(desc(Total))
arrest_by_Date <- (chic_crime[chic_crime$Arrest == 'True',]) %>% group_by(Date) %>% summarise(Total = n())
by_arrest <- chic_crime %>% group_by(Arrest) %>% summarise(Total = n()) %>% arrange(desc(Total))
by_arrest<- na.omit(by_arrest)
crime_by_Date <- chic_crime %>% group_by(Date) %>% summarise(Total = n())
crime_by_Date<-na.omit(crime_by_Date)
by_location <- chic_crime %>% group_by(Location.Description) %>% summarise(Total = n()) %>% arrange(desc(Total))
by_type <- chic_crime %>% group_by(Primary.Type) %>% summarise(Total = n()) %>% arrange(desc(Total))
crime_count_month <- chic_crime %>% group_by(Month) %>% summarise(Total = n())
crime_count_month<- na.omit(crime_count_month)
months <- c("Jan","Feb","Mar","Apr","May","Jun","Jul","Aug","Sep","Oct","Nov","Dec")
crime_count_month$Month <- months[ crime_count_month$Month ]
crime_count_week <- chic_crime %>% group_by(Weekday) %>% summarise(Total = n())
crime_count_week<- na.omit(crime_count_week)
weekdays <- c("Sunday","Monday","Tuesday","Wednesday","Thursday","Friday","Saturday")
crime_count_week$Weekday <- weekdays[ crime_count_week$Weekday ]
crime_count_day <- chic_crime %>% group_by(Day) %>% summarise(Total = n())
crime_count_day<- na.omit(crime_count_day)
theft <- chic_crime[chic_crime$Primary.Type=="THEFT",]
chic_crime_theft <-  theft %>% group_by(Year, Primary.Type) %>% summarise(Total = n())
chic_crime_theft<- na.omit(chic_crime_theft)
battery <- chic_crime[chic_crime$Primary.Type=="BATTERY",]
chic_crime_battery <-  battery %>% group_by(Year, Primary.Type) %>% summarise(Total = n())
chic_crime_battery<- na.omit(chic_crime_battery)
criminal <- chic_crime[chic_crime$Primary.Type=="CRIMINAL DAMAGE",]
chic_crime_criminal <-  criminal %>% group_by(Year, Primary.Type) %>% summarise(Total = n())
chic_crime_criminal<- na.omit(chic_crime_criminal)
narcotics <- chic_crime[chic_crime$Primary.Type=="NARCOTICS",]
chic_crime_narcotics <-  narcotics %>% group_by(Year, Primary.Type) %>% summarise(Total = n())
chic_crime_narcotics<- na.omit(chic_crime_narcotics)
chic_crime_type <- rbind(chic_crime_theft, chic_crime_battery, chic_crime_criminal, chic_crime_narcotics)
streets <- chic_crime[chic_crime$Location.Description=="STREET",]
chic_crime_street <-  streets %>% group_by(Year, Location.Description) %>% summarise(Total = n())
chic_crime_street<- na.omit(chic_crime_street)
residence <- chic_crime[chic_crime$Location.Description=="RESIDENCE",]
chic_crime_residence <-  residence %>% group_by(Year, Location.Description) %>% summarise(Total = n())
chic_crime_residence<- na.omit(chic_crime_residence)
apt <- chic_crime[chic_crime$Location.Description=="APARTMENT",]
chic_crime_apt <-  apt %>% group_by(Year, Location.Description) %>% summarise(Total = n())
chic_crime_apt<- na.omit(chic_crime_apt)
sidewalk <- chic_crime[chic_crime$Location.Description=="SIDEWALK",]
chic_crime_sidewalk <-  sidewalk %>% group_by(Year, Location.Description) %>% summarise(Total = n())
chic_crime_sidewalk<- na.omit(chic_crime_sidewalk)
chic_crime_loc <- rbind(chic_crime_street, chic_crime_residence, chic_crime_apt, chic_crime_sidewalk)
homicide <- chic_crime[chic_crime$Primary.Type=="HOMICIDE",] 
homicide_by_year <-  homicide %>% group_by(Year) %>% summarise(Total = n())
```

## Creating Time Series
```{r}
date_tseries <- xts(by_Date$Total, order.by=as.POSIXct(by_Date$Date))
arrest_tseries <- xts(arrest_by_Date$Total, order.by=as.POSIXct(by_Date$Date))
```

# Plotting the Time Series Graph

## Plot for Crimes and Arrest since 2008-2016
```{r}
hchart(date_tseries, name = "Crimes") %>% 
  hc_add_series(arrest_tseries, name = "Arrests") %>%
  hc_add_theme(hc_theme_flat()) %>%
  hc_title(text = "Times Series plot of Chicago Crimes and Arrests [2008 -2016]") %>%
  hc_legend(enabled = TRUE)
```
  
## Plot for Arrests made since 2008-2016
```{r}
hchart(arrest_tseries) %>%
  hc_add_theme(hc_theme_economist()) %>%
  hc_title(text = "Times Series plot of Arrests made in Chicago [2008 -2016]")
```
  
## Crimes by Year and Month since 2008-2016
```{r}
crime_count <- chic_crime %>% group_by(Year, Month) %>% summarise(Total = n())
crime_count<- na.omit(crime_count)
crime <- ggplot(crime_count, aes(Year, Month, fill = Total)) +
  geom_tile(size = 1, color = "white") +
  theme_classic() +
  geom_text(aes(label=Total), color='white') +
  ggtitle("Crimes by Year and Month[2008-2016]")
plot(crime)
```
  
## Arrests by Year and Month since 2008-2016
```{r}
arrest_data <- (chic_crime[chic_crime$Arrest == 'True',])
arrest_count <- arrest_data %>% group_by(Year, Month) %>% summarise(Total = n())
arrest_count<- na.omit(arrest_count)
arrest <- ggplot(arrest_count, aes(Year, Month, fill = Total)) +
  geom_tile(size = 1, color = "white") +
  theme_classic() +
  scale_fill_viridis()  +
  geom_text(aes(label=Total), color='white') +
  ggtitle("Arrests by Year and Month [2008-2016]")
plot(arrest)
```

## Types of Crimes since 2008-2016
```{r}
hchart(by_type[1:10,],'column', hcaes(x = Primary.Type, y = Total, color = Total)) %>%
  hc_add_theme(hc_theme_flat()) %>%
  hc_plotOptions(column = list(stacking = 'normal')) %>%
  hc_legend(align = 'right', float = T)%>%
  hc_title(text = "Top 10 types of crime")
```

## Major Crime Types since 2008-2016
```{r}
hchart(by_type[1:10,], 'pie', hcaes(x = Primary.Type, y = Total, color = Total)) %>%
  hc_add_theme(hc_theme_google()) %>%
  hc_title(text = "Top 10 major crime") 
```

```{r}
hchart(chic_crime_type,'line', hcaes(x = Year, y = Total, group = Primary.Type)) %>%
  hc_add_theme(hc_theme_google()) %>%
  hc_plotOptions(column = list(stacking = 'normal')) %>%
  hc_legend(align = 'right', float = T)%>%
  hc_title(text = "Major Crime Types")
```
  
## Number of crimes by month of the year
```{r}
hchart(crime_count_month,'column', hcaes(x = Month, y = Total, color = Total)) %>%
  hc_plotOptions(column = list(stacking = 'normal')) %>%
  hc_legend(align = 'right', float = T)%>%
  hc_title(text = "Number of crimes by month of the year")
```

## Number of crimes by Day of the Week
```{r}
hchart(crime_count_week,'column', hcaes(x = Weekday, y = Total, color = Total)) %>%
  hc_plotOptions(column = list(stacking = 'normal')) %>%
  hc_legend(align = 'right', float = T)%>%
  hc_title(text = "Number of crimes by Day of the Week")
```

## Number of crimes by Day of the year
```{r}
hchart(crime_count_day,'column', hcaes(x = Day, y = Total, color = Total)) %>%
  hc_plotOptions(column = list(stacking = 'normal')) %>%
  hc_legend(align = 'right', float = T)%>%
  hc_title(text = "Number of crimes by Day of the year")
```

## Crime count since 2008-2016
```{r}
crime_count <- chic_crime %>% group_by(Year) %>% summarise(Total = n())
crime_count<- na.omit(crime_count)
hchart(crime_count, "column", hcaes(Year,Total, color = Year)) %>%
  hc_add_theme(hc_theme_gridlight()) %>%
  hc_title(text = "Crime Count 2008-2016")
```
  
## Arrest count since 2008-2016
```{r}
arrest_count <- arrest_data %>% group_by(Year) %>% summarise(Total = n())
arrest_count<- na.omit(arrest_count)
hchart(arrest_count, "column", hcaes(Year,Total, color = Year)) %>%
  hc_add_theme(hc_theme_gridlight()) %>%
  hc_title(text = "Arrest Count 2008-2016") 
```
 
## Types of Crime Location since 2008-2016
```{r}
hchart(by_location[1:10,], "column", hcaes(x = Location.Description, y = Total, color = Total)) %>%
  hc_add_theme(hc_theme_flat()) %>%
  hc_title(text = "Top 10 Crime Locations") %>%
  hc_legend(enabled = FALSE)
```

## Major Crime Locations by Year

```{r}
hchart(chic_crime_loc,'column', hcaes(x = Year, y = Total, group = Location.Description)) %>%
  hc_add_theme(hc_theme_google()) %>%
  hc_plotOptions(column = list(stacking = 'normal')) %>%
  hc_legend(align = 'right', float = T)%>%
  hc_title(text = "Major Crime Locations")
```

## Homicides in Chicago since 2008-2016
```{r}
hchart(homicide_by_year, "column", hcaes(Year, Total, color = Year)) %>%
  hc_add_theme(hc_theme_flat()) %>%
   hc_plotOptions(column = list(stacking = 'normal')) %>%
  hc_legend(align = 'right', float = T)%>%
  hc_title(text = "Homicide since 2008-2016")
```

```{r}
homicide_count <- homicide %>% group_by(Year, Month) %>% summarise(Total = n())
ggplot(homicide_count, aes(Year, Month, fill = Total)) +
  geom_tile(size = 1, color = "white") +
  theme_classic() +
  geom_text(aes(label=Total), color='white') +
  ggtitle("Homicides in Chicago since 2008-2016")
```

# Forecasting using Facebook's Prophet Package

## Forecasting Crime in Chicago
```{r}
names(crime_by_Date) <- c("ds","y")
df <- crime_by_Date %>%
  mutate(y = log(y))
m <- prophet(df)
future <- make_future_dataframe(m, periods = 365)
forecast <- predict(m, future)
plot(m, forecast,ylab="log(Total_Crime)",xlab="Year")
prophet_plot_components(m, forecast)
```

# For Random Forest and kNNN Algorithm
## Preparing the dataset
```{r}
chic_crime <- read.table("D:/Study/Northeastern/Fall-2017/DM-ML/Homework/Project/Dataset/Chicago_Crimes_2012_to_2016.csv",header=TRUE,sep=",",fill=TRUE,colClasses = c('NULL',NA,NA,NA,NA,NA,NA,NA,NA,NA,NA,NA,NA,NA,NA,NA,NA,NA,NA,NA,NA,NA,NA))
chic_crime = as.data.table(chic_crime)
names(chic_crime) = tolower(names(chic_crime))
chic_crime = chic_crime[year>2011&year<2017]
chic_crime = chic_crime[complete.cases(chic_crime),]
setkey(chic_crime,id)
```

## Creating Train Data
```{r}
ind_train = seq(1,dim(chic_crime)[1],by=10)
train = chic_crime[ind_train]
train = setkey(train,id)
```

## Creating Test Data
```{r}
ind_test = seq(2,length(ind_train),by=2)
ind_test = ind_train[ind_test]-1
test = chic_crime[ind_test]
test = setkey(test,id)
```

## Training Data for Random Forest
```{r}
train[,cnt := 1]
train[,month := substring(date,1,2)]
train[,year := substring(date,7,10)]
train[,time := substring(date,12,22)]
train[,time := ifelse(grepl("PM",time),as.integer(substring(time,1,2))+12,as.integer(substring(time,1,2)))]
train[,simpledate := paste(month,"01",year,sep="/")]
train[,day := weekdays(as.Date(substring(date,1,10),format="%m/%d/%Y"))]
sum_loc = train[,.(count =sum(cnt)),by=.(location.description)]
cat_loc = head(sum_loc[order(count,decreasing=TRUE),],25)
cat_loc = as.data.frame(cat_loc)
cat_loc = as.character(cat_loc[,1])
sum_type = train[,.(count = sum(cnt)),by=.(primary.type)]
cat_type = head(sum_type[order(count,decreasing=TRUE),],25)
cat_type = as.data.frame(cat_type)
cat_type = as.character(cat_type[,1])
train[,grouplocation := ifelse(location.description %in% cat_loc,as.character(location.description),'Non-Primary')]
train[,groupprimtype := ifelse(primary.type %in% cat_type,as.character(primary.type),'Other')]
```


## Random Forest
```{r}
train_RF = train
setkey(train_RF,id)
train_RF$district = as.factor(train_RF$district)
train_RF$grouplocation = as.factor(train_RF$grouplocation)
train_RF$fbi.code = as.factor(train_RF$fbi.code)
train_RF$ward = as.factor(train_RF$ward)
train_RF$year = as.factor(train_RF$year)
train_RF$time = as.factor(train_RF$time)
train_RF$day = as.factor(train_RF$day)
chic_crime_RF = randomForest(arrest ~ district + ward + grouplocation + year + day + fbi.code + time,mtry=2,ntree=1000,data=train_RF)
chic_crime_RF
```

## k-Nearest Neighbors
# Train and Test data for kNN
```{r}
train_KNN = train[,c("arrest","beat","district","ward","community.area")]
train_KNN$arrest = as.logical(train_KNN$arrest)
test_KNN = test[,c("arrest","beat","district","ward","community.area")]
test_KNN$arrest = as.logical(test_KNN$arrest)
```

## Performing k-Nearest Neighbors
```{r}
testauto_KNN = knn(train = train_KNN, cl = train_KNN$arrest, test = test_KNN, k = 3)
trainauto_KNN = knn(train = train_KNN, cl = train_KNN$arrest, test = train_KNN, k = 3)
table(test_KNN$arrest,testauto_KNN)
table(train_KNN$arrest,trainauto_KNN)
```